Uncertainty propagation for variable (predictor) effects and bias correction in generalized linear (mixed) models???

### 2021-10-20 variable_predictions.Rnw
Autocommit
Temp for focal_pop

When visually presented, predictor effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the predictors whose estimates are usually, but not always, a subject to less clarity of interpretation. Further, generating predictor effects together with the associated confidence intervals for regression models has a number of challenges. In particular: 
\begin{enumerate}
\item choice of representative values of \emph{focal} predictor(s) and appropriate ``model center" for \emph{non-focal} predictors especially in multivariate models
\item  propagation of uncertainty ([how] can we incorporate uncertainty in nonlinear components of (G)LMMs? Should we exclude variation in non-focal parameters?)
\end{enumerate}

Most common way of dealing with the first challenge is taking unique levels of the focal predictor if discrete or taking appropriately sized quantiles (or bins) if continuous, and then calculating the predictions while holding non-focal predictors at their typical values (e.g., averages). This generates -- \emph{predictor effect} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} and \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to this quantity as \emph{predictor effect} since it should, for example, tell us what we would expect the presponse to be at a particular value or level of the predictor. An alternative to averaging the non-focal predictors, is the population-based approach which involves computing the prediction over the population of the non-focal predictors and then averaging across the values of the focal predictor. Dealing with second challenge is discussed in Section~\ref{sec:uncertainity_propagation}

@@

We consider more complex interactions in which the focal predictor is modelled as a cubic polynomial. Specifically:
%
\begin{align}\label{sim:lm_cubic}
\mathrm{hh~size} &= \beta_0 + \beta_{\mathrm{A}}\mathrm{Age}^3_i + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i \nonumber\\
\mathrm{Age}_i &\sim \mathrm{Uniform}(-1, 1) \nonumber\\
\mathrm{Wealthindex}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\epsilon_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\beta_0 &= 1.5 \nonumber\\
\beta_{\mathrm{A}} &= 0.3 \nonumber\\
\beta_{\mathrm{W}} &= 2 \nonumber\\
i &= 1,\cdots, 100
\end{align}
%

>>>>

When visually presented, predictor effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the predictors whose estimates are usually, but not always, a subject to less clarity of interpretation. Further, generating predictor effects together with the associated confidence intervals for regression models has a number of challenges. In particular: 
\begin{enumerate}
\item choice of representative values of \emph{focal} predictor(s) and appropriate ``model center" for \emph{non-focal} predictors especially in multivariate models
\item  propagation of uncertainty ([how] can we incorporate uncertainty in nonlinear components of (G)LMMs? Should we exclude variation in non-focal parameters?)
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)M)
\end{enumerate}

Most common way of dealing with the first challenge is taking unique levels of the focal predictor if discrete or taking appropriately sized quantiles (or bins) if continuous, and then calculating the predictions while holding non-focal predictors at their typical values (e.g., averages). This generates -- \emph{predictor effect} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} and \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to this quantity as \emph{predictor effect} since it should, for example, tell us what we would expect the presponse to be at a particular value or level of the predictor. Formerly, predictor effects computes the expected outcome by meaningfully holding the non-focal predictors constant (or averaged in some meaningful way) while varying the focal predictor, with the goal that the outcome expected prediction represents how the model responds to the changes in the focal predictor.

An alternative to averaging the non-focal predictors, is the population-based approach which involves computing the prediction over the population of the non-focal predictors and then averaging across the values of the focal predictor. Dealing with second challenge is discussed in Section~\ref{sec:uncertainity_propagation}

@@

We consider more complex interactions in which the focal predictor is modelled as a cubic polynomial. Specifically:
%
\begin{align}\label{sim:lm_cubic}
\mathrm{hh~size} &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i \nonumber\\
\mathrm{Age}_i &\sim \mathrm{Uniform}(-1, 1) \nonumber\\
\mathrm{Wealthindex}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\epsilon_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\beta_0 &= 1.5 \nonumber\\
\beta_{\mathrm{A}} &= 0.3 \nonumber\\
\beta_{\mathrm{W}} &= 2 \nonumber\\
i &= 1,\cdots, 100
\end{align}
%
